From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-23 upi.tex
outline for paper
my intent for section 2

@@

By restricting to finite types, two operational semantics emerge:
\begin{enumerate}
\item a (reversible) operational semantics of programs over finite types, and
\item a (reversible) operational semantics of program transformers of such programs.
\end{enumerate}
Such programs can be interpreted as generalized permutations, as they work on
structure as well as the underlying skeleton of finite types, aka the natural numbers.
Many of the transformers can be \emph{oriented}, and thus be interpreted as
\emph{optimizations} in some sense.

In this paper we will
identify a ``featherweight'' representative to use for the formal
development. But first, we illustrate the flavor of the family of languages with
some examples written in a Haskell syntax. First consider the following program:

@@

%%%%%
\subsection{$\Pi_{\mathbb{2}}$}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}

>>>>

POPL 2012: Information effects; Canonicity for 2-dimensional type theory; look
different but we show are the ``same''.

@@

By restricting to finite types, two operational semantics emerge:
\begin{enumerate}
\item a (reversible) operational semantics of programs over finite types, and
\item a (reversible) operational semantics of program transformers of such programs.
\end{enumerate}
$\Pi$ programs can be interpreted as generalized permutations, as they work on
structure as well as the underlying skeleton of finite types, aka the natural numbers.
Many of the transformers can be \emph{oriented}, and thus be interpreted as
\emph{optimizations}.  As with traditional languages, there are different aspects
we can optimize for (program size, number of steps to completion, etc); we will
not touch on this further in this paper.

We will later identify a ``featherweight'' version of $\Pi$ to use in our formal
development.  But first we illustrate the flavor of the family of languages with
some examples.

\jacques{Example 1: using the combinators from pi-dual, some an equational
derivation in type theory (which?), in equational-derivation style.  Point out
that reading ``down'' the justifications is a $\Pi$ program which is also a
constructive proof of the equivalence of the types.  As it is reversible, we
can also read the reverse program by reading ``up''.}

\jacques{Example 2: like example 1, but this time at the next level (i.e. a
program equivalence).  Same remark old.  Refer to Shulmann notes, and point out
that such equational proofs at this level can actually be interpreted in appropriate
categories as commutative diagrams.}

\jacques{Refer to some Sabry et al. papers regarding Toffoli gates.}

\begin{comment}
In this paper we will
identify a ``featherweight'' representative to use for the formal
development. But first, we illustrate the flavor of the family of languages with
some examples written in a Haskell syntax. First consider the following program:

@@

\end{comment}
%%%%%
\subsection{$\Pi_{\mathbb{2}}$}

@@

Canonical forms: given a level 1 program p, the level 2 combinators are
complete; they can transform p to either id or not:
\begin{verbatim}
cmpl2-lem : (p : B <->1 B) -> (p <->2 id) + (p <->2 not)
\end{verbatim}

@@

That space {2} is a higher groupoid. Let's look at its objects; its 1-paths; its
2-paths etc.

Show Agda code

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Correspondence}

Soundness; completeness; etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
~
\newpage
\section{Old}
